---
title: Scare-off measures against wild boars 
subtitle: Computational Movement Analysis
author: Lia Baumann and Michele Senn
output: html_document
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}

#set the setting for all code chunks:

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load wild boar data and libraries, include=FALSE}
#load packages
library(devtools)
#devtools::install_github("ComputationalMovementAnalysis/ComputationalMovementAnalysisData")
library(ComputationalMovementAnalysisData)
library(ggplot2)
library(tidyverse)
library(sf)
library(zoo)
library(ggpubr)
```

# Abstract
This project aims to develop a method for modelling the behavior of wild boars who have been in contact with scare-off measures. The hypothesis is that their behavior can be modeled by their distance to the corresponding Schreck, as well as comparing the average time they spent in a certain field before the scare-off measure called "Schreck" was put into action, during and after.

As a result, the distance of three boars named Caroline, Olga and Sabine to a Schreck in the region Fanel (WSS-2015-01) was visualized. The model depicts also the Schreck activity, which changed from standard to aggressive mode, and the moments when the boars spent time inside of the bean field where the Schreck was located in.
Additionally, the relative change in distance was shown as a second order effect.
The behavior could be modelled by distance in a simple way but it is still lacking validation and the amount of data might not have been enough to draw conclusions.

# Introduction
This project aims to develop a method for modelling the behavior of wild boars who have been in contact with scare-off measures. 

## Research Question
Once a scare-off measure has been set into action, it needs to be examined weather or not it has an effect. Therefore, the following research question was formulated:
**How can the behavior of the boars who had been in contact with a Schreck be modelled?**

The hypothesis is that their behavior can be modeled by their distance to the corresponding Schreck, as well as comparing the average time they spent in a certain field before the scare-off measure called "Schreck" was put into action, during and after.

Additionally, a second question was formulated, which was used to limit the scope of the research question:
**How can a return pattern to the location of the Schreck be modelled?**

## Data
Data about the "Schrecks" and their activity agenda, about the boars movement tracking and field data about the placement of the Schreck was used in this project.

## Analytical concepts
Spatial and time overlaps between the tracked animals were plotted to get an idea of what the data looks like and finding patterns like which individual is affected by which Schreck.
Distances to Scare-Off-Location over time were calculated using the st_distance() function, which returns the shortest distance that separates two geometries.
The function st_intersects() provides intersection of geographical data points.
With the Rolling Window method rollmean(), sampling size was reduced and compared to a sampling size reduction by omitting observations.
To find returning patterns by building a "In the Field / not in the Field" vs. distance to Schreck graph, the application of joins and disjoints and appending a column with the binary data was implemented.

## R concepts
The modelling was done using the following concepts and packages:
Spatial objects were created and manipulated with the r-package "sf".
Raster data and background maps were plotted and visualized with these r-packages: "ggplot2"
For the reshaping and transforming of data the r-package "tidyverse" was used.
In the "zoo" package, the function rollmean() is contained, and arrangement of plots was partially done by using ggarrange() from "ggpubr".

## Risk analysis
The scope is considered a risk because of the amount of behaviors that could potentially be analysed. It is therefore advised to limit the scope. It was done by using a secondary research question.
The Interpretation of behaviors might be vague or false because an emerging pattern is only correlated but not causal.
Technical difficulties could also pose a risk, e.g. when handling too much data in the pre-processing stage, choosing irrelevant thresholds or interpretation issues (semantic annotations).
Possibly, the different measures don't show the same effects. It might be necessary to distinguish between measures in order to draw inferences.

Disclaimer:
Data preparation and conceptualization was divided between the two team members.
The contributors were marked where possible.

# 2. Methods

## 2.1 Data pre-processing
The data was loaded and converted into the coordination system crs = 2056. To find out, which boars were affected by which Schreck, the datasets were both transformed into the same coordinate system using st_transform. They were subsequently plotted on top of each other to visualize the spatial overlap. *(Lia)* 

```{r temporal overlap, eval=FALSE, include=FALSE}
#load data and put into readable format
sampling_periods <- wildschwein_BE %>%
  group_by(TierID, TierName, CollarID) %>%
  summarise(
    min = min(DatetimeUTC),
    max = max(DatetimeUTC)
  )


wildschwein_overlap_temp <- wildschwein_overlap_temp %>%
  left_join(sampling_periods, by = c("TierID", "TierName", "CollarID"))


wildschwein_overlap_temp %>%
mutate(TierCollar = paste(TierName, CollarID)) %>%
  ggplot(aes(xmin = min, xmax = max, y = TierCollar)) +
  geom_errorbarh() +
  facet_grid(Groups~., scales = "free_y", space = "free_y")

```

```{r spatial overlap, include=FALSE}
#convert into same coordinate system

wildschwein_sf <- wildschwein_BE %>%
  st_as_sf(coords = c("E", "N"), crs = 2056) %>%
  mutate(tiercollar = paste(TierID, TierName, CollarID)) 

wildschwein_convex_hull <- wildschwein_sf %>%
  group_by(TierID, TierName, CollarID) %>%
  summarise() %>%
  st_convex_hull()


wildschwein_convex_hull %>%
  mutate(tiercollar = paste(TierID, TierName, CollarID)) %>%
  ggplot(aes(fill = factor(TierID))) + geom_sf(alpha = 0.1) +
  coord_sf(datum = 2056) +
  facet_wrap(~tiercollar) +
  theme(legend.position = "none")


st_overlaps(wildschwein_convex_hull, sparse = FALSE)[1:5, 1:5]
```

```{r EDA}
#looking at all data points

#schreck_locations needs to be transformed into the same coordinate system
schreck_locations_sf <- schreck_locations %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(2056)

#plot wildschweins and schrecks together to see their spatial overlap
ggplot() +
  geom_sf(data=wildschwein_sf,aes(fill=TierID)) +
  geom_sf(data=schreck_locations_sf, aes(color=region)) +
  coord_sf(datum=2056)
```

Apparently, the animals were only tracked around the Schreck in the region of Fanel. As a consequence, all other Schreck data points were removed and the spatial overlap replotted to validate the finding. *(Lia)*

```{r filter all but region Fanel}
#only Fanel
schreck_locations_sf_fanel <- schreck_locations_sf %>%
  filter(region=="fanel")

ggplot() +
  geom_sf(data=wildschwein_sf,aes(fill=TierID)) +
  geom_sf(data=schreck_locations_sf_fanel, aes(color=region)) +
  coord_sf(datum=2056) + theme_classic()
```

In order to combine the data about the Schreck locations and their corresponding agenda, or sounding time, they were joined by their matching column "id" with an inner join.
Following this join, the agenda of all the "Schreck time" was visualized using an errorbar plot.
Table 1 shows a table of the possibly relevant Schrecks as they were active at some point during the boar tracking time. *(Lia)*

```{r join schreck agendas fanel, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
agenda_fanel <- inner_join(schreck_locations_sf_fanel, schreck_agenda, by="id")

#show when the schrecks in fanel were used
ggplot(agenda_fanel, aes(xmin = datum_on, xmax = datum_off, y=id )) +
  geom_errorbarh()

knitr::kable(agenda_fanel %>%
               filter(lubridate::year(datum_off)<2017) %>%
               select(id, datum_on, datum_off, modus),
             caption = "Table 1: All Schrecks in region Fanel during boar tracking time"
             )
```

## 2.2 Selection of focus
In order to see which animals were tracked during the active period of the "Schreck", the boar data was filtered by the dates in which the Schreck were active.
For the tracked boars, a distance between their coordinates and the Schreck coordinate was calculated using st_distance(). In the following graph, one example of the resulting distances is shown. *(Lia)*

```{r distance to Schreck 1}
#1 Schreck filter
schreck1 <- agenda_fanel[1,]

#filter for Schreck period

wildschwein_schreck1 <- filter(wildschwein_sf,DatetimeUTC>schreck1$datum_on,DatetimeUTC<schreck1$datum_off)

#add st_distance as a column
wildschwein_schreck1$distance <- as.numeric(st_distance(wildschwein_schreck1,schreck1)[,1])

#ggplot

ggplot(wildschwein_schreck1, aes(DatetimeUTC,distance, color=TierName)) + geom_point() + labs(title="Distance of the affected boars to Schreck WSS_2014_04") + theme_classic() + labs(x="",y="Distance to WSS_2014_04 [m]")
```


Since it isn't known how close you would need to be to hear the Schreck, all of the five boars Caroline, Isabelle, Nicole, Sabine and Ueli could potentially have been affected.

Subsequenty, the same procedure was repeated for the other Schrecks in the Fanel region (results not displayed). *(Lia)*

```{r distance to Schreck 2, include=FALSE}
#filter for schreck and running period
schreck2 <- agenda_fanel[2,]

wildschwein_schreck2 <- filter(wildschwein_sf,DatetimeUTC>schreck2$datum_on,DatetimeUTC<schreck2$datum_off)

#add distance
wildschwein_schreck2$distance <- as.numeric(st_distance(wildschwein_schreck2,schreck2)[,1])
#ggplot

ggplot(wildschwein_schreck2, aes(DatetimeUTC,distance, color=TierName)) + geom_point() + labs(title="Distance of the affected boars to Schreck WSS_2014_05") + theme_classic()
```

```{r distance to Schreck 5, include=FALSE}
#filter
schreck5 <- agenda_fanel[5,]

wildschwein_schreck5 <- filter(wildschwein_sf,DatetimeUTC>schreck5$datum_on,DatetimeUTC<schreck5$datum_off)

#add distance
wildschwein_schreck5$distance <- as.numeric(st_distance(wildschwein_schreck5,schreck5)[,1])

#ggplot

ggplot(wildschwein_schreck5, aes(DatetimeUTC,distance, color=TierName)) + geom_point() + labs(title="Distance of the affected boars to Schreck WSS_2015_01") + theme_classic()
```

```{r distance to Schreck 6, include=FALSE}
#filter
schreck6 <- agenda_fanel[6,]

wildschwein_sf$distance <- as.numeric(st_distance(wildschwein_sf,schreck6)[,1])

affected_distance6 <- wildschwein_sf %>%
  filter(TierName == c("Caroline","Olga","Sabine"), DatetimeUTC > "2015-05-01 21:01:14",DatetimeUTC < "2015-08-12 21:01:14")

#ggplot
ggplot(affected_distance6, aes(DatetimeUTC,distance, color=TierName)) + geom_point() + labs(title="Distance of the affected boars to Schreck WSS_2015_01") + theme_classic()
```


```{r distance to Schreck 10, eval=FALSE, include=FALSE}
#filter
schreck10 <- agenda_fanel[10,]

wildschwein_schreck10 <- filter(wildschwein_sf,DatetimeUTC>schreck10$datum_on,DatetimeUTC<schreck10$datum_off)

#add distance
wildschwein_schreck10$distance <- as.numeric(st_distance(wildschwein_schreck10,schreck10)[,1])

#ggplot

ggplot(wildschwein_schreck10, aes(DatetimeUTC,distance, color=TierName)) + geom_point() + labs(title="Distance of the affected boars to Schreck WSS_2015_04")
```

The analysis led to the conclusion that only one Schreck WSS_2015_04, first with standard mode and subsequently switching to aggressive mode, can provide enough data to be used for the modelling steps.
**Sabine, Olga and Caroline** were tracked during the running time, and also before and after. *(Lia)*

All following conceptual steps were carried out with this focus.


## 2.3 Movement conceptualization: Distance to Schreck
Using the approach of calculating the distance of the boar to the Schreck plotted against the time provides insights about the movement radius around the Schreck during inactive and active times. *(Lia)*

```{r Caroline and Schreck 5}
distance5_longerThanSchrecking <- wildschwein_sf %>%
  filter(TierName == "Caroline")
distance5_longerThanSchrecking$distance_Schreck5 <- as.numeric(st_distance(distance5_longerThanSchrecking,schreck5)[,1])

ggplot(distance5_longerThanSchrecking,aes(DatetimeUTC,distance_Schreck5)) + geom_point() + labs(y ="Distance to WSS-2015-01 [m]", x="",title="Caroline's distance during whole tracking time") + theme_classic()
```


Zooming in for better understanding of the pattern:

```{r Caroline zoom in}

distance5_4days_Caroline <- wildschwein_sf %>%
  filter(TierName == "Caroline", DatetimeUTC > "2015-04-01", DatetimeUTC < "2015-04-05")
distance5_4days_Caroline$distance_Schreck5 <- as.numeric(st_distance(distance5_4days_Caroline,schreck5)[,1])

ggplot(distance5_4days_Caroline,aes(DatetimeUTC,distance_Schreck5)) + geom_point() + labs(y ="Distance to WSS-2015-01 [m]", x="",title="Caroline's distance during 4 days") + theme_classic()
```


Caroline's moving pattern shows her staying in one place during the days (mostly between 6-9 am to around 6pm). Speculation: That's most likely when she's sleeping. Movements, or better to say: changing of distances, are happening only at night, which was to be expected of a nocturnal animal. *(Lia)*


## 2.4 Relating locations to environment and defining field
As a next step, finding which kind of field this Schreck WSS-2015-01 is in, was done with the additional vector dataset "Feldaufnahmen_Fanel.gpkg" which provides geographical vector data about the agricultural fields in Fanel.
The intersection of Schreck and corresponding field was located using "st_intersects" from the sf-package. *(Lia)*


```{r Schreck 5 and 6 is in a bean field, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
forest <- read_sf("Feldaufnahmen_Fanel.gpkg")
schreck5 <- agenda_fanel[5,]
schreck6 <- agenda_fanel[6,]
#schreck9 <- agenda_fanel[9,]
schreck10 <- agenda_fanel[10,]

st_intersects(schreck5, forest) # sagt, wo sich der Punkt des Schrecks mit den Feld-Daten kreuzt
#Resultat: 272 -->FieldID 195
#st_intersects(schreck6,forest) #same field

#st_intersects(schreck10,forest)

forest_bohnen195 <- forest %>%
  filter(FieldID==195)

ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=schreck5) + labs(title="Intersection of WSS-2015-01 and field", subtitle= "Schreck is in a bean field") + theme_classic()
```

The following graphs show Caroline's location around the bean field during active WSS-2015-01, first during standard mode and then during aggressive mode. These graphs seem to have an issue with scaling which makes it difficult to compare the spread of locations. *(Lia)*

```{r Caroline during Schreck 5 and 6}
#5+6 ist in einem Feld, wo Caroline rumgeht

distance5_Caroline_Schrecktime <- wildschwein_sf %>%
  filter(TierName == "Caroline", DatetimeUTC > "2015-05-20", DatetimeUTC < "2015-06-12")
distance6_Caroline_Schrecktime <- wildschwein_sf %>%
  filter(TierName == "Caroline", DatetimeUTC > "2015-06-12", DatetimeUTC < "2015-07-01")

ggarrange(
  ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=distance5_Caroline_Schrecktime) + labs(title="Standard mode") + theme_classic(),
  ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht))+
    geom_sf(data=distance6_Caroline_Schrecktime) + labs(title="Aggressive mode") + theme_classic(),
  legend="none", widths=c(1,2), align="v"
)

```


Looking at only the bean field:

```{r Caroline zoom in active Schreck 5_6, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
intersection_195andSchreck5 <- st_intersects(forest_bohnen195,distance5_Caroline_Schrecktime)
intersection_195andSchreck6 <- st_intersects(forest_bohnen195,distance6_Caroline_Schrecktime)
#class(intersection_195andSchreck5)

inter_sel_5 <- distance5_Caroline_Schrecktime[intersection_195andSchreck5[[1]],]
inter_sel_6 <- distance6_Caroline_Schrecktime[intersection_195andSchreck6[[1]],]

ggarrange(
  ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_5) +
    geom_sf(data=schreck5, aes(size=5)) + labs(title="Standard mode") + theme_classic(),
  ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_6) +
    geom_sf(data=schreck6, aes(size=5)) + labs(title="Aggressive mode") + theme_classic(),
  legend="none"
)
  #during Schreck time, Carolina still visited the bean field (during Standard and aggressive mode)
```


As can be seen, WSS-2015-01 is in a bean field (mode standard and then aggressive). Caroline made some visits during Schreck time.

In a similar way, the positions outside the bean field could be found by subsetting the dataset with st_disjoint() from the sf-package. *(Lia)*

```{r Caroline in the field vs not in the field Schreck 5 and 6, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#ich will jetzt das Feld definieren. Im Feld drin = TRUE --> alle in inter_sel_5
#für die "nicht enthaltenen" kann ich einen anti-join nutzen oder das Subsetting mit disjoint

disjoint_195andSchreck5 <- st_disjoint(forest_bohnen195,distance5_Caroline_Schrecktime)
disjoint_195andSchreck6 <- st_disjoint(forest_bohnen195,distance6_Caroline_Schrecktime)
#class(intersection_195andSchreck5)

disjoint_caroline_5 <- distance5_Caroline_Schrecktime[disjoint_195andSchreck5[[1]],]
disjoint_caroline_6 <- distance6_Caroline_Schrecktime[disjoint_195andSchreck6[[1]],]

#disjoint anschauen
ggarrange(
  ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+
    geom_sf(data=disjoint_caroline_5) +
    labs(title="Outside field during standard mode"),
  ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+
    geom_sf(data=disjoint_caroline_6) +
    labs(title="Outside during aggressive mode"),
  legend="none"
)
```

Using the information about the inside and outside positions, respectively, and adding binary information (1= "in field", 0= "outside of field"), creates information about when Caroline was moving inside of the bean field and when not. The following graphs depict this concept by showing red points where Caroline was inside of the bean field. *(Lia)*

```{r Caroline in and out}
#nun die ausserhalb mit 0 und die innerhalb mit 1 ergänzen
disjoint_caroline_5$inField <- 0
disjoint_caroline_6$inField <- 0
inter_sel_5$inField <- 1
inter_sel_6$inField <- 1
#append the two data sets again with their new columns
Caroline_Schreck5_withField <- bind_rows(disjoint_caroline_5,inter_sel_5)
Caroline_Schreck6_withField <- bind_rows(disjoint_caroline_6,inter_sel_6)
#bis hier stimmt es schon mal

Caroline_Schreck5_withField$distance5 <- as.numeric(st_distance(Caroline_Schreck5_withField$geometry,schreck5)[,1])
Caroline_Schreck6_withField$distance6 <- as.numeric(st_distance(Caroline_Schreck6_withField$geometry,schreck6)[,1])

#this is now the pattern during Schreck time --> red points don't say a lot yet, just that she was in the field --> conceptualize better?
ggarrange(
  ggplot() + geom_point(data=Caroline_Schreck5_withField,aes(DatetimeUTC,distance5)) +
  geom_point(data=inter_sel_5,aes(DatetimeUTC,inField, colour="in Field")) + labs(title="Standard mode", y="Distance to WSS-2015-01 [m]", x="", colour="Location") + ylim(c(0,2100)),
  ggplot() + geom_point(data=Caroline_Schreck6_withField,aes(DatetimeUTC,distance6)) +
  geom_point(data=inter_sel_6,aes(DatetimeUTC,inField, colour="in Field")) + labs(title="Aggressive mode", y="", x="", colour="Location"),
  common.legend=TRUE, legend="bottom"
) ##scaling of y axis not ok yet
```

To show the distances or movements, respectively, not only during active Schreck times, but also before and after, the whole tracking time of Caroline was visualized. *(Lia)*

```{r Caroline longer time span, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
Caroline_longerThanSchrecktime <- wildschwein_sf %>%
  filter(TierName == "Caroline")

Caroline_longerThanSchrecktime$distance5_6 <-
  as.numeric(st_distance(Caroline_longerThanSchrecktime$geometry,schreck5)[,1])
#don't need to do it for Schreck 6 because it is the same with different mode

#ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=Caroline_longerThanSchrecktime)

intersection_195andSchreck5_6_long <- st_intersects(forest_bohnen195,Caroline_longerThanSchrecktime)
#class(intersection_195andSchreck5)

inter_sel_5_6_long <- Caroline_longerThanSchrecktime[intersection_195andSchreck5_6_long[[1]],]

#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_5_6_long) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Caroline's positions in the bean field during the whole time")
  #during Schreck time, Carolina still visited the bean field

#ich will jetzt das Feld definieren. Im Feld drin = TRUE --> alle in inter_sel_5_long
#für die "nicht enthaltenen" kann ich einen anti-join nutzen oder das Subsetting mit disjoint

disjoint_195andSchreck5_6_long <- st_disjoint(forest_bohnen195,Caroline_longerThanSchrecktime)
#class(intersection_195andSchreck5_6_long)

disjoint_caroline_5_6_long <- Caroline_longerThanSchrecktime[disjoint_195andSchreck5_6_long[[1]],]

#disjoint anschauen --> ok!
#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=disjoint_caroline_5_6_long) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Caroline's positions outside bean field during the whole time")

#nun die ausserhalb mit 0 und die innerhalb mit 1 ergänzen
disjoint_caroline_5_6_long$inField <- 0
inter_sel_5_6_long$inField <- 1
#append the two data sets again with their new columns
Caroline_Schreck5_6_withField_long <- bind_rows(disjoint_caroline_5_6_long,inter_sel_5_6_long)

#this is now the pattern during the whole time --> red points are in Field
ggplot() + geom_point(data=Caroline_Schreck5_6_withField_long,aes(DatetimeUTC,distance5_6)) +
  geom_point(data=inter_sel_5_6_long,aes(DatetimeUTC,inField, colour="red"), show.legend=FALSE) +
  labs(title="Caroline's distance to WSS-2015-01 during whole tracking time",subtitle="Red points: in the bean field", x="",y="Distance to WSS-2015-01 [m]")
#hier noch  zu ergänzen: Linien, wann die Schrecks aktiv sind (z.B. https://stackoverflow.com/questions/66825661/vertical-lines-for-data-points-in-geom-point-in-ggplot2-for-r)
```

Reduction of the timescale to March 2015 until November 2015 for Caroline: *(Lia)*

```{r Caroline smaller timescale}
Caroline_Schreck5_6_withField_long_aroundInFieldtime <- Caroline_Schreck5_6_withField_long %>%
  filter(DatetimeUTC>"2015-04-15", DatetimeUTC<"2015-08-15")
inter_sel_5_6_long_shorter <- inter_sel_5_6_long %>%
  filter(DatetimeUTC<"2015-08-15")

ggplot() + geom_point(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime,aes(DatetimeUTC,distance5_6)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(DatetimeUTC,inField, colour="red"), show.legend=FALSE) +
  labs(title="Caroline's distance to WSS-2015-01 around active Schreck time",subtitle="Red points: in the bean field", x="",y="Distance to WSS-2015-01 [m]")
#hier noch  zu ergänzen: Linien, wann die Schrecks aktiv sind (z.B. https://stackoverflow.com/questions/66825661/vertical-lines-for-data-points-in-geom-point-in-ggplot2-for-r)
```


## 2.5 Displaying  influence of Schreck activity

To conceptualize the active and inactive time of the Schreck, a background fill was used, provided with the function geom_rect() from ggplot2. *(Lia)*


```{r Caroline with background color}
 Caroline_breaks <-
  data.frame(start=as.Date(c("2015-04-15","2015-05-20","2015-06-12","2015-07-01")),
             end=as.Date(c("2015-05-20","2015-06-12","2015-07-01","2015-08-15")),
             colors=factor(1:4),
             name=c("Before Schreck activity","During active Schreck 5","During active Schreck 6","After Schreck activity"))

Caroline_breaks$Phases <- factor(Caroline_breaks$name, as.character(Caroline_breaks$name))

ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) +
  geom_point(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime,aes(as.Date(DatetimeUTC),distance5_6)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(as.Date(DatetimeUTC),inField, colour="red"), show.legend=FALSE) +
  labs(title="Caroline's distance to WSS-2015-01 around active Schreck time",subtitle="Red points: in the bean field", x="",y="Distance to WSS-2015-01 [m]", colour="")

#Variante mit weniger Datenpunkten:
Caroline_Schreck5_6_withField_long_aroundInFieldtime_smooth <- Caroline_Schreck5_6_withField_long_aroundInFieldtime %>%
  dplyr::filter(row_number() %% 150 == 0)
Caroline_Schreck5_6_withField_long_aroundInFieldtime_smooth2 <- Caroline_Schreck5_6_withField_long_aroundInFieldtime %>%
  dplyr::filter(row_number() %% 50 == 0)
```

## 2.6 Evaluating complexity reduction

Subsequently, it was tested if complexity could be reduced by reduction of sampling size to every 50th and 150th observation (choice based on visual inspection). As a methodological comparison, smoothing was computed using the Rolling Window approach with a k of 10, 100 and 150. *(Lia)*

```{r sampling size reduction Caroline}
#plotten mit sampling size reduction
ggarrange(
  ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime_smooth2,aes(as.Date(DatetimeUTC),distance5_6)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(subtitle= "Reducing sampling size to every 50th observation", colour="Location") +
    rremove("y.title") + rremove("x.title"),
  ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime_smooth,aes(as.Date(DatetimeUTC),distance5_6)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(subtitle= "Reducing sampling size to every 150th observation", colour="Location") +
    rremove("y.title") + rremove("x.title"),
  nrow=2, common.legend=TRUE, legend="right"
)
```


```{r rolling window Caroline}
#Variante mit rolling window:
Caroline_Schreck5_6_withField_long_aroundInFieldtime$rollmean <- rollmean(Caroline_Schreck5_6_withField_long_aroundInFieldtime$distance5_6,2, NA, align="left")
Caroline_Schreck5_6_withField_long_aroundInFieldtime$rollmean2 <- rollmean(Caroline_Schreck5_6_withField_long_aroundInFieldtime$distance5_6,10, NA, align="left")
Caroline_Schreck5_6_withField_long_aroundInFieldtime$rollmean3 <- rollmean(Caroline_Schreck5_6_withField_long_aroundInFieldtime$distance5_6,100, NA, align="left")
Caroline_Schreck5_6_withField_long_aroundInFieldtime$rollmean4 <- rollmean(Caroline_Schreck5_6_withField_long_aroundInFieldtime$distance5_6,150, NA, align="left")

#plotten mit rolling window
ggarrange(
  ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime,aes(as.Date(DatetimeUTC),rollmean2)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(subtitle="Rolling Mean, k = 10", colour="Location") +
    rremove("y.title") + rremove("x.title"),
  ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime,aes(as.Date(DatetimeUTC),rollmean3)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(subtitle="Rolling Mean, k = 100", colour="Location") +
    rremove("y.title") + rremove("x.title"),
  ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime,aes(as.Date(DatetimeUTC),rollmean4)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(subtitle="Rolling Mean, k = 150", colour="Location") +
    rremove("y.title") + rremove("x.title"),
  nrow=3, common.legend=TRUE, legend="right"
)
```

## 2.7 Extension of analyses to more boars

Olga's and Sabine's data were prepared following the same steps as was done with Caroline (preparation steps not shown). *(Lia)*

```{r Olga during Schrecktime 5 and 6}
distance5_Olga_Schrecktime <- wildschwein_sf %>%
  filter(TierName == "Olga", DatetimeUTC > "2015-05-20", DatetimeUTC < "2015-06-12")

distance6_Olga_Schrecktime <-wildschwein_sf %>%
  filter(TierName == "Olga", DatetimeUTC > "2015-06-12", DatetimeUTC < "2015-07-01")

#ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=distance5_Olga_Schrecktime)

#ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=distance6_Olga_Schrecktime)
#Olga was in and around the fields 5 & 6 a lot!

intersection_195andSchreck5_olga <- st_intersects(forest_bohnen195,distance5_Olga_Schrecktime)
intersection_195andSchreck6_olga <- st_intersects(forest_bohnen195,distance6_Olga_Schrecktime)

inter_sel_5_olga <- distance5_Olga_Schrecktime[intersection_195andSchreck5_olga[[1]],]
inter_sel_6_olga <- distance6_Olga_Schrecktime[intersection_195andSchreck6_olga[[1]],]

#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_5_olga) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Olga's positions in the bean field while Schreck 5 was active")
  #during Schreck time, Olga still visited the bean field

#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_6_olga) + geom_sf(data=schreck6, aes(size=5)) + labs(title="Olga's positions in the bean field while Schreck 6 was active")
```


```{r Olga longer time span}
Olga_longerThanSchrecktime <- wildschwein_sf %>%
  filter(TierName == "Olga")

Olga_longerThanSchrecktime$distance5_6 <-
  as.numeric(st_distance(Olga_longerThanSchrecktime$geometry,schreck5)[,1])
#don't need to do it for Schreck 6 because it is the same with different mode

#ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=Olga_longerThanSchrecktime)

intersection_195andSchreck5_6_long_olga <- st_intersects(forest_bohnen195,Olga_longerThanSchrecktime)
#class(intersection_195andSchreck5)

inter_sel_5_6_long_olga <- Olga_longerThanSchrecktime[intersection_195andSchreck5_6_long_olga[[1]],]

#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_5_6_long_olga) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Olga's positions in the bean field during the whole time")
  #during Schreck time, Olga still visited the bean field

#ich will jetzt das Feld definieren. Im Feld drin = TRUE --> alle in inter_sel_5_long
#für die "nicht enthaltenen" kann ich einen anti-join nutzen oder das Subsetting mit disjoint

disjoint_195andSchreck5_6_long_olga <- st_disjoint(forest_bohnen195,Olga_longerThanSchrecktime)
#class(intersection_195andSchreck5_long)

disjoint_Olga_5_6_long <- Olga_longerThanSchrecktime[disjoint_195andSchreck5_6_long_olga[[1]],]

#disjoint anschauen --> ok!
#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=disjoint_Olga_5_6_long) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Olga's positions outside bean field during the whole time")

#nun die ausserhalb mit 0 und die innerhalb mit 1 ergänzen
disjoint_Olga_5_6_long$inField <- 0
inter_sel_5_6_long_olga$inField <- 1
#append the two data sets again with their new columns
Olga_Schreck5_6_withField_long <- bind_rows(disjoint_Olga_5_6_long,inter_sel_5_6_long_olga)

#this is now the pattern during the whole time --> red points are in Field
#ggplot() + geom_point(data=Olga_Schreck5_6_withField_long,aes(DatetimeUTC,distance5_6)) +
#  geom_point(data=inter_sel_5_6_long_olga,aes(DatetimeUTC,inField, colour="red")) +
#  labs(title="Olga's distance to Schreck 5 / 6 during whole tracking time. Red points mean: in the bean field")
#hier noch  zu ergänzen: Linien, wann die Schrecks aktiv sind (z.B. https://stackoverflow.com/questions/66825661/vertical-lines-for-data-points-in-geom-point-in-ggplot2-for-r)
```

```{r Olga smaller timescale}
Olga_Schreck5_6_withField_long_aroundInFieldtime <- Olga_Schreck5_6_withField_long %>%
  filter(DatetimeUTC<"2015-08-01")

#ggplot() + geom_point(data=Olga_Schreck5_6_withField_long_aroundInFieldtime,aes(DatetimeUTC,distance5_6)) +
#  geom_point(data=inter_sel_5_6_long_olga,aes(DatetimeUTC,inField, colour="red")) +
#  labs(title="Olga's distance to Schreck 5 / 6 around inField time. Red points mean: in the bean field")
#hier noch  zu ergänzen: Linien, wann die Schrecks aktiv sind (z.B. https://stackoverflow.com/questions/66825661/vertical-lines-for-data-points-in-geom-point-in-ggplot2-for-r)
```

```{r Olga with background color}
Olga_breaks <- data.frame(start=as.Date(c("2015-04-15","2015-05-20","2015-06-12","2015-07-01")),
                              end=as.Date(c("2015-05-20","2015-06-12","2015-07-01","2015-08-15")),
                              colors=factor(1:4),
                              name=c("Before Schreck activity","During active Schreck 5","During active Schreck 6","After Schreck activity"))

Olga_breaks$Phases <- factor(Olga_breaks$name, as.character(Olga_breaks$name))

#ggplot() +
#  geom_rect(data = Olga_breaks,
 #           aes(xmin = start,
  #              xmax = end,
   #             ymin = - Inf,
  #             ymax = Inf,
  #              fill = colors),
  #          alpha = 0.5) +
  #geom_point(data=Olga_Schreck5_6_withField_long_aroundInFieldtime,aes(as.Date(DatetimeUTC),distance5_6)) +
  #geom_point(data=inter_sel_5_6_long_olga,aes(as.Date(DatetimeUTC),inField, colour="red")) +
  #labs(title="Olga's distance to Schreck 5 / 6 around inField time. Red points mean: in the bean field", colour="")

#Variante mit weniger Datenpunkten:
Olga_Schreck5_6_withField_long_aroundInFieldtime_smooth <- Olga_Schreck5_6_withField_long_aroundInFieldtime %>%
  dplyr::filter(row_number() %% 150 == 0)

#ggplot() +
#  geom_rect(data = Olga_breaks,
 #           aes(xmin = start,
  #              xmax = end,
  #              ymin = - Inf,
   #             ymax = Inf,
  #              fill = Phases),
   #         alpha = 0.5) + 
  #geom_line(data=Olga_Schreck5_6_withField_long_aroundInFieldtime_smooth,aes(as.Date(DatetimeUTC),distance5_6)) +
  #geom_point(data=inter_sel_5_6_long_olga,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
 # labs(title="Olga's distance to Schreck 5 and 6", colour="Location", x="",y="Distance [m]")

```

```{r Sabine during Schrecktime 5 and 6}
distance5_Sabine_Schrecktime <- wildschwein_sf %>%
  filter(TierName == "Sabine", DatetimeUTC > "2015-05-20", DatetimeUTC < "2015-06-12")

distance6_Sabine_Schrecktime <-wildschwein_sf %>%
  filter(TierName == "Sabine", DatetimeUTC > "2015-06-12", DatetimeUTC < "2015-07-01")

#ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=distance5_Sabine_Schrecktime)

#ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=distance6_Sabine_Schrecktime)
#Sabine was in and around the fields 5 & 6 a lot!

intersection_195andSchreck5_Sabine <- st_intersects(forest_bohnen195,distance5_Sabine_Schrecktime)
intersection_195andSchreck6_Sabine <- st_intersects(forest_bohnen195,distance6_Sabine_Schrecktime)

inter_sel_5_Sabine <- distance5_Sabine_Schrecktime[intersection_195andSchreck5_Sabine[[1]],]
inter_sel_6_Sabine <- distance6_Sabine_Schrecktime[intersection_195andSchreck6_Sabine[[1]],]

#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_5_Sabine) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Sabine's positions in the bean field while Schreck 5 was active")
  #during Schreck time, Sabine still visited the bean field

#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_6_Sabine) + geom_sf(data=schreck6, aes(size=5)) + labs(title="Sabine's positions in the bean field while Schreck 6 was active")
```

```{r Sabine longer time span}
Sabine_longerThanSchrecktime <- wildschwein_sf %>%
  filter(TierName == "Sabine")

Sabine_longerThanSchrecktime$distance5_6 <-
  as.numeric(st_distance(Sabine_longerThanSchrecktime$geometry,schreck5)[,1])
#don't need to do it for Schreck 6 because it is the same with different mode

#ggplot(forest_bohnen195)+geom_sf(aes(fill=Frucht)) + geom_sf(data=Sabine_longerThanSchrecktime)

intersection_195andSchreck5_6_long_Sabine <- st_intersects(forest_bohnen195,Sabine_longerThanSchrecktime)
#class(intersection_195andSchreck5)

inter_sel_5_6_long_Sabine <- Sabine_longerThanSchrecktime[intersection_195andSchreck5_6_long_Sabine[[1]],]

#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=inter_sel_5_6_long_Sabine) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Sabine's positions in the bean field during the whole time")
  #during Schreck time, Sabine still visited the bean field

#ich will jetzt das Feld definieren. Im Feld drin = TRUE --> alle in inter_sel_5_long
#für die "nicht enthaltenen" kann ich einen anti-join nutzen oder das Subsetting mit disjoint

disjoint_195andSchreck5_6_long_Sabine <- st_disjoint(forest_bohnen195,Sabine_longerThanSchrecktime)
#class(intersection_195andSchreck5_long)

disjoint_Sabine_5_6_long <- Sabine_longerThanSchrecktime[disjoint_195andSchreck5_6_long_Sabine[[1]],]

#disjoint anschauen --> ok!
#ggplot(forest_bohnen195) + geom_sf(aes(fill=Frucht))+ geom_sf(data=disjoint_Sabine_5_6_long) + geom_sf(data=schreck5, aes(size=5)) + labs(title="Sabine's positions outside bean field during the whole time")

#nun die ausserhalb mit 0 und die innerhalb mit 1 ergänzen
disjoint_Sabine_5_6_long$inField <- 0
inter_sel_5_6_long_Sabine$inField <- 1
#append the two data sets again with their new columns
Sabine_Schreck5_6_withField_long <- bind_rows(disjoint_Sabine_5_6_long,inter_sel_5_6_long_Sabine)

#this is now the pattern during the whole time --> red points are in Field
#ggplot() + geom_point(data=Sabine_Schreck5_6_withField_long,aes(DatetimeUTC,distance5_6)) +
#  geom_point(data=inter_sel_5_6_long_Sabine,aes(DatetimeUTC,inField, colour="red")) +
 # labs(title="Sabine's distance to Schreck 5 / 6 during whole tracking time. Red points mean: in the bean field")
#hier noch  zu ergänzen: Linien, wann die Schrecks aktiv sind (z.B. https://stackoverflow.com/questions/66825661/vertical-lines-for-data-points-in-geom-point-in-ggplot2-for-r)
```


```{r Sabine smaller timescale}
Sabine_Schreck5_6_withField_long_aroundInFieldtime <- Sabine_Schreck5_6_withField_long %>%
  filter(DatetimeUTC>"2015-06-01")

#ggplot() + geom_point(data=Sabine_Schreck5_6_withField_long_aroundInFieldtime,aes(DatetimeUTC,distance5_6)) +
#  geom_point(data=inter_sel_5_6_long_Sabine,aes(DatetimeUTC,inField, colour="red")) +
#  labs(title="Sabine's distance to Schreck 5 / 6 around inField time. Red points mean: in the bean field")
#hier noch  zu ergänzen: Linien, wann die Schrecks aktiv sind (z.B. https://stackoverflow.com/questions/66825661/vertical-lines-for-data-points-in-geom-point-in-ggplot2-for-r)
```

```{r Sabine with background color}
Sabine_breaks <- data.frame(start=as.Date(c("2015-04-15","2015-05-20","2015-06-12","2015-07-01")),
                              end=as.Date(c("2015-05-20","2015-06-12","2015-07-01","2015-08-15")),
                              colors=factor(1:4),
                              name=c("Before Schreck activity","During active Schreck 5","During active Schreck 6","After Schreck activity"))

Sabine_breaks$Phases <- factor(Sabine_breaks$name, as.character(Sabine_breaks$name))

#Variante mit weniger Datenpunkten:
Sabine_Schreck5_6_withField_long_aroundInFieldtime_smooth <- Sabine_Schreck5_6_withField_long_aroundInFieldtime %>%
  dplyr::filter(row_number() %% 150 == 0)


#ggplot() +
#  geom_rect(data = Sabine_breaks,
#            aes(xmin = start,
 #               xmax = end,
  #              ymin = - Inf,
   #             ymax = Inf,
    #            fill = Phases),
     #       alpha = 0.5) + 
#  geom_line(data=Sabine_Schreck5_6_withField_long_aroundInFieldtime_smooth,aes(as.Date(DatetimeUTC),distance5_6)) +
#  geom_point(data=inter_sel_5_6_long_Sabine,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
#  labs(title="Sabine's distance to Schreck 5 and 6", colour="Location", x="",y="Distance [m]")

```

Validating the robustness of the methodolical approach was undertaken by visualization and comparison of their different positions around the bean field during active WSS-2015-01. *(Lia)*

```{r compare Olga and Caroline and Sabine during Schreck time}
ggarrange(
  ggplot(forest_bohnen195) + geom_sf()+ geom_sf(data=inter_sel_5) + geom_sf(data=schreck5, aes(colour="Schreck 5", size=2)) +
    labs(title="Caroline Schreck 5"),
  ggplot(forest_bohnen195) + geom_sf()+ geom_sf(data=inter_sel_5_olga) + geom_sf(data=schreck5, aes(colour="Schreck 5", size=2)) +
    labs(title="Olga Schreck 5"),
  ggplot(forest_bohnen195) + geom_sf()+ geom_sf(data=inter_sel_5_Sabine) + geom_sf(data=schreck5, aes(colour="Schreck 5", size=2)) +
    labs(title="Sabine Schreck 5"),
  ggplot(forest_bohnen195) + geom_sf()+ geom_sf(data=inter_sel_6) + geom_sf(data=schreck6, aes(colour="Schreck 6", size=2)) +
    labs(title="Caroline Schreck 6"),
  ggplot(forest_bohnen195) + geom_sf()+ geom_sf(data=inter_sel_6_olga) + geom_sf(data=schreck6, aes(colour="Schreck 6", size=2)) +
    labs(title="Olga Schreck 6"),
  ggplot(forest_bohnen195) + geom_sf()+ geom_sf(data=inter_sel_6_Sabine) + geom_sf(data=schreck6, aes(colour="Schreck 6", size=2)) +
    labs(title="Sabine Schreck 6"),
  ncol=3,nrow=2,legend="none",common.legend=TRUE
)
```

## 2.8 Visualising relative change of distance to Schreck

For completisation of the movement conceptualization, an additional visualisation was computed. What if the distance cannot depict the movement pattern, but the relative change of the distance provides more insight? Following this question, the relative change of distance during the same time period was plotted. It was calculated by grouping average daily and weekly distances and calculating their average change in %. The resulting graphs will be shown in the chapter "Results". *(Lia)*


# 3. Results

The final conceptualization of movement patterns of the three female boars is depicted in the following graph. The red points correspond to the position being inside of the bean field with the Schreck WSS-2015-01. Sampling size reduction to every 150th observation was chosen due to a sufficient resolution. Losing the information in between seems to be an acceptable compromise for the increase in readability. *(Lia)*

```{r compare distances from all 3 animals, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggarrange(
  ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Caroline_Schreck5_6_withField_long_aroundInFieldtime_smooth,aes(as.Date(DatetimeUTC),distance5_6)) +
  geom_point(data=inter_sel_5_6_long_shorter,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(title="Caroline's distance to Schreck 5 and 6", colour="Location", x="",y="Distance [m]"),
  ggplot() +
  geom_rect(data = Olga_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Olga_Schreck5_6_withField_long_aroundInFieldtime_smooth,aes(as.Date(DatetimeUTC),distance5_6)) +
  geom_point(data=inter_sel_5_6_long_olga,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(title="Olga's distance to Schreck 5 and 6", colour="Location", x="",y="Distance [m]"),
  ggplot() +
  geom_rect(data = Sabine_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.5) + 
  geom_line(data=Sabine_Schreck5_6_withField_long_aroundInFieldtime_smooth,aes(as.Date(DatetimeUTC),distance5_6)) +
  geom_point(data=inter_sel_5_6_long_Sabine,aes(as.Date(DatetimeUTC),inField, colour="In bean field")) +
  labs(title="Sabine's distance to Schreck 5 and 6", colour="Location", x="",y="Distance [m]"),
  ncol=1, nrow=3, legend="none",common.legend= TRUE
  )
```



```{r relative distance change}

Relative_breaks <-
  data.frame(start=(c(15,20,23,26)),
             end=(c(20,23,26,32)),
             colors=factor(1:4),
             Phases=c("Before Schreck activity","During active Schreck 5","During active Schreck 6","After Schreck activity"))
 #daily:
  
  distance5_longerThanSchrecking_relChange_day <- Caroline_Schreck5_6_withField_long_aroundInFieldtime %>%
  mutate(day=lubridate::date(DatetimeUTC)) %>%
  group_by(day) %>%
  arrange(desc(day)) %>%
  summarise(avg_dist = mean(distance5_6)) %>%
  mutate(dist_perc= (avg_dist/lead(avg_dist) - 1) * 100) %>%
  mutate(TierName="Caroline")

distance5_longerThanSchrecking_relChange_olga_day <- Olga_Schreck5_6_withField_long_aroundInFieldtime %>%
  mutate(day=lubridate::date(DatetimeUTC)) %>%
  group_by(day) %>%
  arrange(desc(day)) %>%
  summarise(avg_dist = mean(distance5_6)) %>%
  mutate(dist_perc= (avg_dist/lead(avg_dist) - 1) * 100) %>%
  mutate(TierName = "Olga")

distance5_longerThanSchrecking_relChange_Sabine_day <- Sabine_Schreck5_6_withField_long_aroundInFieldtime %>%
  mutate(day=lubridate::date(DatetimeUTC)) %>%
  group_by(day) %>%
  arrange(desc(day)) %>%
  summarise(avg_dist = mean(distance5_6)) %>%
  mutate(dist_perc= (avg_dist/lead(avg_dist) - 1) * 100) %>%
  mutate(TierName = "Sabine")

relChange_all_3boars_day <- rbind(distance5_longerThanSchrecking_relChange_day,distance5_longerThanSchrecking_relChange_olga_day,distance5_longerThanSchrecking_relChange_Sabine_day)

#weekly:
distance5_longerThanSchrecking_relChange_week <- Caroline_Schreck5_6_withField_long_aroundInFieldtime %>%
  mutate(Week=lubridate::week(DatetimeUTC)) %>%
  group_by(Week) %>%
  arrange(desc(Week)) %>%
  summarise(avg_dist = mean(distance5_6)) %>%
  mutate(dist_perc= (avg_dist/lead(avg_dist) - 1) * 100) %>%
  mutate(TierName="Caroline")

distance5_longerThanSchrecking_relChange_olga_week <- Olga_Schreck5_6_withField_long_aroundInFieldtime %>%
  mutate(Week=lubridate::week(DatetimeUTC)) %>%
  group_by(Week) %>%
  arrange(desc(Week)) %>%
  summarise(avg_dist = mean(distance5_6)) %>%
  mutate(dist_perc= (avg_dist/lead(avg_dist) - 1) * 100) %>%
  mutate(TierName = "Olga")

distance5_longerThanSchrecking_relChange_Sabine_week <- Sabine_Schreck5_6_withField_long_aroundInFieldtime %>%
  mutate(Week=lubridate::week(DatetimeUTC)) %>%
  group_by(Week) %>%
  arrange(desc(Week)) %>%
  summarise(avg_dist = mean(distance5_6)) %>%
  mutate(dist_perc= (avg_dist/lead(avg_dist) - 1) * 100) %>%
  mutate(TierName = "Sabine")

relChange_all_3boars_week <- rbind(distance5_longerThanSchrecking_relChange_week,distance5_longerThanSchrecking_relChange_olga_week,distance5_longerThanSchrecking_relChange_Sabine_week)

ggplot() +
  geom_rect(data = Caroline_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.2) +
  geom_line(data=relChange_all_3boars_day, aes(day,dist_perc)) + labs(y="Change in distance [%]", title="Daily relative change in distance from WSS-2015-01", x="") + theme_classic() +
  facet_grid(TierName~.)

#geom_vline(xintercept=20, linetype="dashed"),

ggplot() +
  geom_rect(data = Relative_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Phases),
            alpha = 0.2) +
  geom_line(data=relChange_all_3boars_week, aes(Week,dist_perc)) + labs(y="Change in distance [%]", title="Weekly relative change in distance from WSS-2015-01", x="") + theme_classic() +
  facet_grid(TierName~.)

```

## In the field vs. out of the field before, during and after approach
In this approach the time of a boar inside a certain area is compared to the time spent outside of this area. The idea behind this approach was to gather information out of a rather large dataset that is relevant for a farmer. Furthermore, this approach should provide some relative numbers with which a conclusion might be reached as to weather or not the scare-off measures have an effect on the boars. 
A binary approach in form of a inside vs. outside the field was used. The defined area is a certain field. This will be done for the period before the Schreck was put into action, during and after. The percentage of time spent inside the field vs. outside the field can then be compared between the timespan before, during and after.

First the Schreck and boar for this approach had to be selected. It had be Schrecks and boars were several animals were in the vicinity of a Schreck during its running period were there's also enough tracking data of the boars movement from the period before the Schreck was put into action and from the period after the Schreck was active. This was done as described above. Schreck number 5 and 6 respectively were selected and the boars Caroline, Sabine and Olga, also as described in more detail above.

The question which this approach seeks to answer is:
How much time does a boar spend in the field before the Schreck (in %), during and after? *(Michèle)*

Schreck 5 and 6 is in a bean field (as described above). Caroline seems to be spending 0.03 % of her time in that field before the Schreck was put there, 2 % of her time during the Schrecks active period and 0.4 % of her time after the Schrecks active period.

Sabine seems to be spending 0 % of her time in that field before the Schreck was put there, 0.9 % of her time during the Schrecks active period and 3 % of her time after the Schrecks active period.

Olga seems to be spending 0.4 % of her time in that field before the Schreck was put there, 2 % of her time during the Schrecks active period and 0.7 % of her time after the Schrecks active period.

```{r Caroline in the field vs not in the field during SChreck time}
st_intersects(schreck5, forest)

fieldforschreck5 <- forest %>%
  filter(FieldID==195)

Caroline_Schrecktime <- wildschwein_sf %>%
  filter(TierName == "Caroline", DatetimeUTC > "2015-05-20", DatetimeUTC < "2015-06-12")

Caroline_field <- st_disjoint(fieldforschreck5,Caroline_Schrecktime)


Caroline_field_during <- Caroline_Schrecktime[Caroline_field[[1]],]

CarolineIntersect <- st_intersects(fieldforschreck5,Caroline_Schrecktime)

intersect <- Caroline_Schrecktime[CarolineIntersect[[1]],]

#out of field = 0 in field = 1 
Caroline_field_during$inField <- 0
intersect$inField <- 1

#append 
Caroline_field_inout_during <- bind_rows(Caroline_field_during,intersect)

  


#47 from 2207 Data points are in the field
print(47/2207*100)
# during Schreck time she spends 2 % of her time in the field.

```

```{r Caroline in the field vs not in the field before SChreck time}

Caroline_Schrecktime_before <- wildschwein_sf %>%
  filter(TierName == "Caroline", DatetimeUTC < "2015-05-20")

Caroline_field <- st_disjoint(fieldforschreck5,Caroline_Schrecktime_before)

Caroline_field_before <- Caroline_Schrecktime_before[Caroline_field[[1]],]

CarolineIntersectbefore <- st_intersects(fieldforschreck5,Caroline_Schrecktime_before)

intersectbefore <- Caroline_Schrecktime_before[CarolineIntersectbefore[[1]],]

#out of field = 0 in field = 1 
Caroline_field_before$inFieldbefore <- 0
intersectbefore$inFieldbefore <- 1

#append 
Caroline_field_inout_before <- bind_rows(Caroline_field_before,intersectbefore)

  


#4 from 14789 Data points are in the field
#print(4/14789*100)
# during Schreck time she spends 0.03 % of her time in the field.

```

```{r Caroline in the field vs not in the field after SChreck time}

Caroline_Schrecktime_after <- wildschwein_sf %>%
  filter(TierName == "Caroline", DatetimeUTC > "2015-06-12")

Caroline_field <- st_disjoint(fieldforschreck5,Caroline_Schrecktime_after)

Caroline_field_after <- Caroline_Schrecktime_after[Caroline_field[[1]],]

CarolineIntersectafter <- st_intersects(fieldforschreck5,Caroline_Schrecktime_after)

intersectafter <- Caroline_Schrecktime_after[CarolineIntersectafter[[1]],]

#out of field = 0 in field = 1 
Caroline_field_after$inField <- 0
intersectafter$inField <- 1

#append 
Caroline_field_inout_after <- bind_rows(Caroline_field_after,intersectafter)

  


#156 from 39029 Data points are in the field
#print(156/39029*100)
# during Schreck time she spends 0.4 % of her time in the field.

```

```{r Sabine in the field vs not in the field during SChreck time}

Sabine_Schrecktime <- wildschwein_sf %>%
  filter(TierName == "Sabine", DatetimeUTC > "2015-05-20", DatetimeUTC < "2015-06-12")

Sabine_field <- st_disjoint(fieldforschreck5,Sabine_Schrecktime)


Sabine_field_during <- Sabine_Schrecktime[Sabine_field[[1]],]

SabineIntersect <- st_intersects(fieldforschreck5,Sabine_Schrecktime)

intersect <- Sabine_Schrecktime[SabineIntersect[[1]],]

#out of field = 0 in field = 1 
Sabine_field_during$inField <- 0
intersect$inField <- 1

#append 
Sabine_field_inout_during <- bind_rows(Sabine_field_during,intersect)

  


#19 from 2207 Data points are in the field
print(19/2207*100)
# during Schreck time she spends 0.9 % of her time in the field.

```

```{r Sabine in the field vs not in the field before SChreck time}

Sabine_Schrecktime_before <- wildschwein_sf %>%
  filter(TierName == "Sabine", DatetimeUTC < "2015-05-20")

Sabine_field <- st_disjoint(fieldforschreck5,Sabine_Schrecktime_before)

Sabine_field_before <- Sabine_Schrecktime_before[Sabine_field[[1]],]

SabineIntersectbefore <- st_intersects(fieldforschreck5,Sabine_Schrecktime_before)

intersectbefore <- Sabine_Schrecktime_before[SabineIntersectbefore[[1]],]

#out of field = 0 in field = 1 
Sabine_field_before$inField <- 0
intersectbefore$inField <- 1

#append 
Sabine_field_inout_before <- bind_rows(Sabine_field_before,intersectbefore)

  


#0 from 16165 Data points are in the field
#print(0/16165*100)
# during Schreck time she spends 0 % of her time in the field.

```

```{r Sabine in the field vs not in the field after SChreck time}

Sabine_Schrecktime_after <- wildschwein_sf %>%
  filter(TierName == "Sabine", DatetimeUTC > "2015-06-12")

Sabine_field <- st_disjoint(fieldforschreck5,Sabine_Schrecktime_after)

Sabine_field_after <- Sabine_Schrecktime_after[Sabine_field[[1]],]

SabineIntersectafter <- st_intersects(fieldforschreck5,Sabine_Schrecktime_after)

intersectafter <- Sabine_Schrecktime_after[SabineIntersectafter[[1]],]

#out of field = 0 in field = 1 
Sabine_field_after$inField <- 0
intersectafter$inField <- 1

#append 
Sabine_field_inout_after <- bind_rows(Sabine_field_after,intersectafter)

  


#126 from 4372 Data points are in the field
print(126/4372*100)
# during Schreck time she spends 3 % of her time in the field.

```

```{r Olga in the field vs not in the field during SChreck time}

Olga_Schrecktime <- wildschwein_sf %>%
  filter(TierName == "Olga", DatetimeUTC > "2015-05-20", DatetimeUTC < "2015-06-12")

Olga_field <- st_disjoint(fieldforschreck5,Olga_Schrecktime)


Olga_field_during <- Olga_Schrecktime[Olga_field[[1]],]

OlgaIntersect <- st_intersects(fieldforschreck5,Olga_Schrecktime)

intersect <- Olga_Schrecktime[OlgaIntersect[[1]],]

#out of field = 0 in field = 1 
Olga_field_during$inField <- 0
intersect$inField <- 1

#append 
Olga_field_inout_during <- bind_rows(Olga_field_during,intersect)

  


#42 from 2205 Data points are in the field
#print(42/2205*100)
# during Schreck time she spends 2 % of her time in the field.

```

```{r Olga in the field vs not in the field before SChreck time}

Olga_Schrecktime_before <- wildschwein_sf %>%
  filter(TierName == "Olga", DatetimeUTC < "2015-05-20")

Olga_field <- st_disjoint(fieldforschreck5,Olga_Schrecktime_before)

Olga_field_before <- Olga_Schrecktime_before[Olga_field[[1]],]

OlgaIntersectbefore <- st_intersects(fieldforschreck5,Olga_Schrecktime_before)

intersectbefore <- Olga_Schrecktime_before[OlgaIntersectbefore[[1]],]

#out of field = 0 in field = 1 
Olga_field_before$inField <- 0
intersectbefore$inField <- 1

#append 
Olga_field_inout_before <- bind_rows(Olga_field_before,intersectbefore)

  


#7 from 1921 Data points are in the field
#print(7/1921*100)
# during Schreck time she spends 0.4 % of her time in the field.

```

```{r Olga in the field vs not in the field after SChreck time}

Olga_Schrecktime_after <- wildschwein_sf %>%
  filter(TierName == "Olga", DatetimeUTC > "2015-06-12")

Olga_field <- st_disjoint(fieldforschreck5,Olga_Schrecktime_after)

Olga_field_after <- Olga_Schrecktime_after[Olga_field[[1]],]

OlgaIntersectafter <- st_intersects(fieldforschreck5,Olga_Schrecktime_after)

intersectafter <- Olga_Schrecktime_after[OlgaIntersectafter[[1]],]

#out of field = 0 in field = 1 
Olga_field_after$inField <- 0
intersectafter$inField <- 1

#append 
Olga_field_inout_after <- bind_rows(Olga_field_after,intersectafter)

  


#140 from 19826 Data points are in the field
#print(140/19826*100)
# during Schreck time she spends 0.7 % of her time in the field.

```

```{r plot beforeduringafter in field percentage}
# beforeduringafter <- data.frame(Caroline = c ("before"= 0.03, "during"= 2, "after"= 0.4), Sabine = c("before"= 0, "during"= 0.9, "after"= 0.4), Olga = c("before"= 0.4, "during"= 2, "after"= 0.7))

beforeduringafter <- data.frame(before = c ("Caroline"= 0.03, "Sabine"= 0, "Olga"= 0.4), during = c(2, 0.9, 2), after = c(0.4, 3, 0.7))

dd <- cbind(beforeduringafter$before,beforeduringafter$during, beforeduringafter$after)
barplot(dd, beside=T,  names.arg=c('before','during','after'), legend = c("Caroline", "Sabine", "Olga"),xlim = c(0, 21), ylab = "time spent in field (%)")

```

In order to validate this their wherabouts were plotted before, during and after the Schrecks active period. To visualize weather or not they were inside the field in which the Schreck was placed, the area of the field is marked in the plot.

```{r Caroline beforeduringafter plot inoutside field}
#Caroline
ggarrange(
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Caroline_Schrecktime_before, size = 0.2) + labs(title="Caroline before Schrecktime") + theme_classic(), 
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Caroline_Schrecktime, size = 0.2) + labs(title="Caroline during Schrecktime") + theme_classic(), 
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Caroline_Schrecktime_after, size = 0.2) + labs(title="Caroline after Schrecktime") + theme_classic() 
)

```

```{r Sabine beforeduringafter plot inoutside field}
#Sabine
ggarrange(
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Sabine_Schrecktime_before, size = 0.2) + labs(title="Sabine before Schrecktime") + theme_classic(), 
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Sabine_Schrecktime, size = 0.2) + labs(title="Sabine during Schrecktime") + theme_classic(), 
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Sabine_Schrecktime_after, size = 0.2) + labs(title="Sabine after Schrecktime") + theme_classic() 
)

```

```{r Olga beforeduringafter plot inoutside field}

#Olga
ggarrange(
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Olga_Schrecktime_before, size = 0.2) + labs(title="Olga before Schrecktime") + theme_classic(), 
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Olga_Schrecktime, size = 0.2) + labs(title="Olga during Schrecktime") + theme_classic(), 
  ggplot(fieldforschreck5)+geom_sf(aes(fill=Frucht)) +
    geom_sf(data=Olga_Schrecktime_after, size = 0.2) + labs(title="Olga after Schrecktime") + theme_classic() 
)

```

# 3. Discussion
## 3.1 Distance to conceptualize movement
As the results show, all boars moved in a 0 to 1500 m radius around the Schreck before the sound started, moved around it in a 0 to 1000 m radius during standard mode and increased radius when switching to aggressive and then quiet mode (= after sounding). Caroline's and Olga's movements could be compared, whereas Sabine moved in a different way.
Distance was chosen to conceptualize movement, when plotting against the time on the x axis. The question is, if it can be considered **useful** in order to conceptualize an actual pattern. Additionally, validity of the relationships might not be general enough. With such a small analysis (3 boars with one Schreck), it could easily capture only a peculiarity of the current data. (2) *(Lia)* 

The above explained topic was assessed by using another way of conceptualizing the distance: looking at the relative change of distance in relation to the last day or week (Idea from (4)). It could be considered as distinguishing between first order and second order effects (3).
Looking at the resulting graphs, it can't be concluded that this measure provides a more useful pattern. At least, the findings from the distance graphs can be reproduced this way (especially the similarity between Olga and Caroline) which could be considered as a validation step. *(Lia)* 

# 3.2 In Field vs. Not In Field
The original concept to annotate the times when the boars were inside of the bean field resulted in a creation of point at "in Field" moments. The binary display of In vs. Not In didn't provide any useful pattern so it was omitted from the project. In the results, red points can be seen and followed, and interestingly, the boars were visiting the field even during Schreck time. Not even the aggressive mode during activity stopped them from returing to the field.
This concept could start the analysis of finding returning patterns, though more validation would need to be executed. *(Lia)* 

## 3.3 Smoothing methods
Counting as an interesting finding was the fact that the boars seemed to reduce and increase their distance on a daily basis, which could relate to their behaviors (sleeping outside of the field and moving back during their active days). This semantic annotation is purely speculative, as it was not further analysed (5).
In order to reduce this cycling data overload, smoothing methods very used to show general patterns that would naturally decrease the degree of granularity (6).
One of these methods was subsetting the data to every 50th or 150th element.
The second option was the rolling window method that uses a mean from different window sizes, for this project set at k=c(10,100,150). The outputs of these two methods are slightly different but all increased readability of the general trend in the distances.
Subsetting and reducing sample size to every 150th observation was used for the result, mainly due to the highest degree of readability. *(Lia)* 

## 3.4 Validation
Validation was carried out derived from options in the Rykiel paper (1). It consisted mainly of replotting data either with a subset of the data or in comparison with other parameters (e.g. IDs oder boars). *(Lia)* 

## 3.5 Problems and possible solutions
- Before actual conceptualization could take place, a lot of time was invested to find the affected boars and which Schrecks they were affected by, a necessary task to continue the work on the research question. Analysing data from another researcher without having accompanied the process means that more pre-processing and exploratory data analysis needs to be done. On the other hand, it might lead to finding unexpected patterns, because no expectations were present.
- At this point, I'm not completely sure if the results are even remotely interesting or if it was merely a result in visualisation instead of pattern detection. Maybe these data don't allow the finding of patterns.
- Our initial research question was about finding returning patterns to the Schreck location. Even though the red points show movement inside the field even during active Schreck times, it might not have been the ideal display. So far, this has not yet been overcome.
- One main issue was the difficulty to estimate whether the animals would hear the Schrecks or not - does it depend on distance or is it more related to topological or environmental factors (e.g. wind, density of forest / field)? *(Lia)*

# Outlook
The project delivers a method to visualize the returning behavior before, during and after Schreck activity in a simplistic way. An open question remains, if the behavior was suitably modelled by distance and change in distance. The following next actions might help deriving more insights on the matter:
- The inclusion of context data is suggested so more meaningful conclusions can be drawn. This would include weather data, maps, topological data, ... (3)
- A possible next step could be segmentation of trajectories so behavior patterns can be classified. Ruling out inactive times might show more distinct patterns or at least serve as validation method. (6)
- Could the displacement as an additional alternative to distance and change in distance provide alternative and more meaningful patterns?

Also, it remains unvalidated if the findings are a pattern or mere peculiarity of the data. The next actions are suggested:
- Data collection with a distinct research question about behavior and returning. During this time, tracking less animals during longer time spans and varying the Schreck activity in function of the boar's vicinity.
- Extend the analysis with the existing data set to other boars and Schrecks (if at all possible).


# References
(1) Rykiel, E. J. (1996). Testing ecological models: the meaning of validation. Ecological modelling, 90(3), 229-244.
(2) T4 – Defining & detecting patterns, Patrick Laube
(3) T5 – Context, context, context, Patrick Laube
(4) Molloy, J., Schatzmann, T., Schoeman, B., Tchervenkov, C., Hintermann, B., & Axhausen, K. W. (2021). Observed impacts of the Covid-19 first wave on travel behaviour in Switzerland based on a large GPS panel. Transport Policy, 104, 43-51.
(5) SHAMOUN-BARANES, J., et al. 2012. From Sensor Data to Animal Behaviour: An Oystercatcher Example. PLoS ONE, 7(5), e37997
(6) Laube, Patrick, and Ross S. Purves. 2011. “How Fast Is a Cow? Cross - Scale Analysis of Movement Data.” Transactions in GIS 15 (3): 401–18. https://doi.org/10.1111/j.1467-9671.2011.01256.x.

```{r info text for Michèle}
# From Patrick's project intro:

#Your report shall:
#– cover how you went about investigating your research questions. Describe your data science ideas
#and how you implemented your ideas. - done
#– present the results of your study and discuss them in the light of your research questions. What
#have you achieved and what would be further steps for future research?
#– report problems and limitations you encountered along the way and the solutions you chose to
#overcome these, be it limitations with respect to the data sources, the tools or any other source of
#limitation.
#– discuss your data science choices in the light of the theory covered in the lectures, group works,
#and your reading assignments.
```
